---
title: "Homework5"
author: "Megan Jasek, Rohan Thakur, Charles Kekeh"
date: "Monday, March 7, 2016"
output: pdf_document
---

```{r,include=FALSE}
library(pastecs)
library(knitr)
opts_chunk$set(tidy.opts=list(width.cutoff=70))
```

Question1
==========
Part 1
------
The package has already been installed.

Part 2
------

```{r, tidy=TRUE}
#Loading the astsa library
library(astsa)
```

Part 3
------

```{r, tidy=TRUE, tidy.opts=list(width.cutoff=70)}
#Examining series structures
str(EQ5)
str(flu)
str(gas)
```

Part 4
------
```{r, tidy=TRUE, tidy.opts=list(width.cutoff=70)}
#Plotting for EQ5
plot(EQ5,type="o", main="Seismic Trace of Earthquake number 5 series",
     ylab="Seismic trace of Earthquake",
     xlab="Time unit", col="blue")
hist(EQ5,main="Simulated White Noise", col="blue",
     xlab="Seismic Trace of Earthquake number 5 series", ylab="Frequency")

#Plotting for flu
plot(flu,type="o", main="Monthly pneumonia and influenza deaths in the U.S., 1968 to 1978",
     ylab="Pneumonia and influenza deaths",
     xlab="Year", col="blue")
hist(flu,main="Monthly pneumonia and influenza deaths in the U.S., 1968 to 1978", col="blue",
     xlab="Pneumonia and influenza deaths", ylab="Frequency")

#Plotting for gas
plot(gas,type="o", main="Weekly Gas Prices",
     ylab="Gas Prices",
     xlab="Year", col="blue")
hist(gas,main="Simulated White Noise", col="blue",
     xlab="Seismic Trace of Earthquake number 5 series", ylab="Frequency")
```



Part 5
------
**EQ5** -  Mean of the series stable around 0. Amplitude/Volatility of the series increases sharply around time unit 1000, and fluctuates.

**flu** - Seems to be a downward trend from the 60's leading into the 70's. After this, the trend remains flat with a steady seasonal variation with some years worse than others.

**gas** - The weekly gas price forms a downward trend from 200-2002, upward trend from 2002 - 2009, a sharp decrease in 2009, and a gradual increase from 2009 onwards.


Question 2
==========
**Example 1** -> Number of users adopting new product feature at social media startup (daily data for 1 year). The graph follows an upward trend for three months, corresponding to the initial promotion of the feature, followed by a flattening of the line.

**Example 2**-> The daily stock price of TSLA in the past 3 months. The price followed a steady downward trend until the past few weeks when it has made an upward trend.

**Example 3** -> Another example.



Question 3
==========
```{r, tidy=TRUE, tidy.opts=list(width.cutoff=70)}
#Simulating white noise
white.noise=rnorm(1000,0,1) 

#Plotting timeplot, histogram
plot.ts(white.noise, main="Simulated White Noise", col="navy",
        ylab="Simulated values", xlab="Simulated Time Period")
hist(white.noise,main="Simulated White Noise", col="blue",
     xlab="Simulated Values")
```


Question 4
==========
Unsure of comments.
```{r, tidy=TRUE, tidy.opts=list(width.cutoff=70)}
#Simulating First series
white.noise.series.1 = rnorm(1000,0,1)
series.1 <- white.noise.series.1
for (t in 2:length(white.noise.series.1)){
  series.1[t] <- 0.9*series.1[t-1] + white.noise.series.1[t]}

#Simulating Second Series
white.noise.series.2 = rnorm(1000,0,1)
series.2 <- white.noise.series.2
for (t in 2:length(white.noise.series.2)){
  series.2[t] <- 0.9*series.2[t-1] + white.noise.series.2[t]}

#Plotting Graphs
par(mfrow=c(2,2))
plot.ts(series.1, main="Simulated AR(ar=c(0.9)) Series", col="navy",
        ylab="Simulated values", xlab="Simulated Time Period")
hist(series.1,main="Simulated AR(ar=c(0.9)) Series", col="blue",
     xlab="Simulated Values")
plot.ts(series.2, main="Simulated AR(ar=c(0.2)) Series", col="navy",
        ylab="Simulated values", xlab="Simulated Time Period")
hist(series.2,main="Simulated AR(ar=c(0.2)) Series", col="blue",
     xlab="Simulated Values")
```


Question 5
==========
Unsure of comments.
```{r, tidy=TRUE, tidy.opts=list(width.cutoff=70)}
white.noise.series.new=rnorm(1000,0,1)
r.walk=cumsum(white.noise.series.new) 

# Random walk with drift = 0.5
noise.drift = 0.5 + white.noise.series.new; 
r.walk.drift = cumsum(noise.drift) 

par(mfrow=c(1,1))
plot.ts(r.walk.drift, main="Random Walk with Drift, Random Walk without Drift, Deterministic Trend", col="blue", ylab="Values", xlab="Simulated Time Period", bg=38)
lines(0.5*(1:length(r.walk.drift))+10, lty="dashed", col="navy")
lines(r.walk, col="purple")
# Add vertical lines
abline(v=c(100,200,300,400),col=3,lty=3)
# Add Legend
leg.txt <- c("RW with Drift", "Deterministic Linear Trend", "RW without Drift")
legend("topleft", legend=leg.txt, lty=c(1,2,1), col=c("blue","navy","purple"),
       bty='n', cex=1, merge = TRUE, bg=336)
par(mfrow=c(2,2))
hist(r.walk.drift, main="RW with Drift", col="blue")
hist(0.5*(1:length(r.walk.drift))+10, main="Deterministic Linear Trend", col="navy")
hist(r.walk, main="RW without Drift", col="purple")
```


