---
title: "Homework1"
author: "Megan Jasek, Charles Kekeh, Rohan, Thakur"
date: "Sunday, January 31, 2016"
output: pdf_document
---

```{r,include=FALSE}
library(ggplot2)
library(car)
library(boot)
library(pastecs)
library(car)
```

Question 1
==========
```{r}
# Load the dataframe
load(file.path("./data", "birthweight_w271.rdata"))
```

Question 2
==========
```{r, tidy=TRUE}
# Display summary information about the dataframe
print(desc)
print(str(data))
print(summary(data))
print(stat.desc(data))
```
There are 1388 obs. of 14 variables in the data.

Question 3
==========
```{r, tidy=TRUE}
# Get summary statistics about the birthweight variable.
print(summary(data$bwght))
print(sum(is.nan(data$bwght)))
print(quantile(data$bwght, probs = c(.01, .05, .10, .25, .50, .75, .90, .95, .99, 1)))

# Plot the histogram of bwght at 15 bins
bwght.hist <- ggplot(data, aes(bwght)) + 
    theme(legend.position = "none") + 
    geom_histogram(fill = "Blue", colour = "Black", binwidth = (range(data$bwght)[2] - range(data$bwght)[1])/15) +
    labs(title = "Distribution of birthweight", x = "Birth Weight", y = "Fequency")

plot(bwght.hist)

# Plot the histogram of bwght at 30 bins
bwght.hist <- ggplot(data, aes(bwght)) + 
    theme(legend.position = "none") + 
    geom_histogram(fill = "Blue", colour = "Black") +
    labs(title = "Distribution of birthweight", x = "Birth Weight", y = "Fequency")

plot(bwght.hist)

# Plot the histogram of bwght at 60 bins
bwght.hist <- ggplot(data, aes(bwght)) + 
    theme(legend.position = "none") + 
    geom_histogram(fill = "Blue", colour = "Black", binwidth = (range(data$bwght)[2] - range(data$bwght)[1])/60) +
    labs(title = "Distribution of birthweight", x = "Birth Weight", y = "Fequency")

plot(bwght.hist)
```

Comments on shape of distributions:
-----------------------------------
As more and more bins are added the shape of the distribution gets smoother.  

Data observations:
------------------
Below are the outlier observations. There are babies with birthweights of 0 and over 200 ounces.
```{r}
print(sum(data$bwght == 0))
print(data$bwght[data$bwght > 200])
```
We should remove the zero baby weights from the data. They probably correspond to data entry issues.
We should also remove the single 271 ounces baby observation in the data set. For the purpose of a linear regression, that outlier data point may affect the regression.

Question 4
==========
```{r, tidy=TRUE}
# Get summary statistics about the cigarettes smoked variable.
print(summary(data$cigs))
print(sum(is.nan(data$cigs)))
print(quantile(data$cigs, probs = c(.01, .05, .10, .25, .50, .75, .90, .95, .99, 1)))

# Plot the histogram of cigs at 15 bins
cigs.hist <- ggplot(data, aes(cigs)) + 
    theme(legend.position = "none") + 
    geom_histogram(fill = "Blue", colour = "Black", binwidth = (range(data$cigs)[2] - range(data$cigs)[1])/15) +
    labs(title = "Distribution of daily cigarette consumption", x = "Daily cigarette consumption", y = "Fequency")

plot(cigs.hist)

# Plot the histogram of cigs at 30 bins
cigs.hist <- ggplot(data, aes(cigs)) + 
    theme(legend.position = "none") + 
    geom_histogram(fill = "Blue", colour = "Black") +
    labs(title = "Distribution of daily cigarette consumption", x = "Daily cigarette consumption", y = "Fequency")

plot(cigs.hist)

# Plot the histogram of cigs at 60 bins
cigs.hist <- ggplot(data, aes(cigs)) + 
    theme(legend.position = "none") + 
    geom_histogram(fill = "Blue", colour = "Black", binwidth = (range(data$cigs)[2] - range(data$cigs)[1])/60) +
    labs(title = "Distribution of daily cigarette consumption", x = "Daily cigarette consumption", y = "Fequency")

plot(cigs.hist)
```

Comments on shape of distributions:
-----------------------------------
As more and more bins are added, more granularity of the data can be seen.  

Data observations:
------------------

The histogram for the number of cigarettes smoked is positively skewed with a very high proportion of individuals smoking zero cigarettes per day during their pregnancy.
There are no other visible signs of anomalies in the data.

Question 5:
===========
```{r, tidy=TRUE}
# Create a scatterplot of cigarettes smoked per day vs baby birthweight
scatter.bwght.cigs <- ggplot(data, aes(cigs, bwght)) +
    geom_point(colour = "Blue", position = "jitter") +
    geom_smooth(method = "lm", colour = "Red") +
    labs(x = "Cigarettes Smoked Per Day", 
         y = "Infant Birthweight", 
         title = "Cigarettes Smoked Per Day Vs Infant Birthweight")
plot(scatter.bwght.cigs)
```
Based on the scatterplot and the fitted lm curve on it, it appears that only a very small amount of the variation of bwght will be explained by cigs. That's because the variation explained in the graph appears to be much lower than the variation of birthweights at any level of daily cigarette consumption in the scatterplot.

Question 6:
===========
```{r, tidy=TRUE}
# Clean up the data based on previous observations
data <- data[data$bwght != 0 & data$bwght < 200, ]

# Now perform the OLS regression
simple.ols.cigs.bwght <- lm(bwght ~ cigs, data = data)
print(summary.lm(simple.ols.cigs.bwght))
plot(simple.ols.cigs.bwght)
```

Coefficients:
-------------
__Intercept:__ estimate = 119.6663, standard error = 0.5645\newline
__cigs:__ estimate = -0.5083, standard error = 0.0889\newline
The intercept and slope coefficents of the model are statistically significant.

Interpret the Results:
----------------------

1.  Our null hypothesis is that there is no relationship between the cigs variable and the bwght variable. (The coefficient for the cigs variable is 0).
2.  We are able to reject the null hypothesis since our p-value of the t-statistic of the cigs variable is significant at 1.32e-08.
3.  A change of 1 unit in cigs corresponds to a 0.51 reduction in birthweight.  (The model shows a negative coefficient for the variable cigs with a value of -0.508).
4.  Practical significance:  we have an R-squared value of 0.02322, indicating that 2.32% of the variation in bwght is explained by our model.  An R value of 0.152 indicates a relatively small effect size.

Question 7
==========
```{r, tidy=TRUE}
# Obtain descriptive statistics for the new variable
print(summary(data$faminc))
print(sum(is.nan(data$faminc)))
print(quantile(data$faminc, probs = c(.01, .05, .10, .25, .50, .75, .90, .95, .99, 1)))
print(stat.desc(data$faminc, basic=FALSE, norm = TRUE))

# Plot the histogram of faminc at 30 bins
faminc.hist <- ggplot(data, aes(faminc)) + 
    theme(legend.position = "none") + 
    geom_histogram(fill = "Blue", colour = "Black") +
    labs(title = "Distribution of Family Income", x = "Family Income", y = "Fequency")

plot(faminc.hist)

# Produce a scatterplot of bwght, cigs and faminc
scatterplotMatrix(~ bwght + cigs + faminc, data = data)
```

Data observations:
------------------

In the famiily income variable, it seems that while all values below 20 were collected as exact values, values above may have been collected as ranges. For example, respondents may have ticked boxes such as 20-25, 25-20, 30-35, etc., and in the final variable it seems the data is represented as the mean of the range.  It also seems strange that 65 is so far above the rest of the values. It seems that values above a certain number have been denoted as 65. While this is not ideal, we will proceed with these values as our observations.

Question 8:
==========
```{r, tidy=TRUE}
#Introduce a new independent variable to the model
multiple.ols.cigs.faminc.bwght <- lm(bwght ~ cigs + faminc, data = data)
print(summary.lm(multiple.ols.cigs.faminc.bwght))
plot(multiple.ols.cigs.faminc.bwght)
```

Coefficients:
-------------
__Intercept:__  estimate = 116.97540, standard error = 1.03242\newline
__cigs:__  estimate = -0.45981, standard error = 0.08998\newline
__faminc:__  estimate = 0.08921, standard error = 0.02870\newline
The intercept and slope coefficents of the model are statistically significant.


Interpret the Results:
----------------------
1.  Our null hypothesis is that there is no relationship between the bwght variable and the cigs and faminc variables. (The coefficients for the cigs variable and faminc variable are 0).
2.  We are able to reject the null hypothesis since our p-value of the t-statistic of the cigs variable is significant at 3.67e-07 and our p-value of the t-statistic of the faminc variable is significant at 0.00192.
3.  A change of 1 unit in cigs corresponds to a 0.46 reduction in birthweight.  (The model shows a negative coefficient for the variable cigs with a value of -0.45981).  A change of 1 unit in faminc corresponds to a 0.09 increase in birthweight.  (The model shows a positive coefficient for the variable faminc with a value of 0.08921).
4.  Practical significance:  we have an R-squared value of 0.03004, indicating that 3.00% of the variation in bwght is explained by our model.  An R value of 0.173 indicates a relatively small effect size.

Question 9
==========
In multiple regression, the coefficient on cigs means that for every additional
cigarette smoked per day by the pregnant mother, leaving the income variable constant,
the birth weight decreases by 0.460 ounces.

In simple regression, we saw that this coefficient was also negative and had a value of -0.508.
In this case, the coefficient on cigs meant that for every additional
cigarette smoked per day, indepedent of any other condition of the mother, the birth
weight was reduced by 0.508 ounces.

In the multiple regression, the variance explained by the faminc variable was captured in the residuals
of the simple model and partially also in the coefficent of the cigs 
variable as we can suspect there is some correlation between the two variables.

Therefore, the introduction of the additional variable faminc has reduced the contribution of the cigs variable to the birth weight. We can hypothesise that there is some correlation between the number of cigarettes smoked per day and the family income, where mothers with higher family income have better health habits and therefore smoke less.
And introducing the family income variable thus takes away some of the variance explanation previously captured
by the cigs variable.

Question 10
===========
The more negative cigs coefficient is that of the simpel model. It's value is -0.508 compared to the -0.46 value for the multiple regression model.  Our explaination for the difference as stated in Question 9 is that there is a correlation between the cigs variable and the faminc variable.