---
title: "Homework4"
author: "Megan Jasek, Rohan Thakur, Charles Kekeh"
date: "Saturday, February 20, 2016"
output: pdf_document
---

```{r,include=FALSE}
library(car)
library(ggplot2)
library(sandwich)
library(lmtest)
library(pastecs)
library(knitr)
opts_chunk$set(tidy.opts=list(width.cutoff=70))
```

```{r, tidy=TRUE, tidy.opts=list(width.cutoff=70)}
# Load the dataset and print descriptions
load("athletics.RData")
desc
```

Question1
==========

How many observations and variables are in the dataset?
-------------------------------------------------------
There are 116 observations of 14 variables in the data set.

Data Analysis
-------------
- __apps__ - The histogram shows a data distribution that's positevely skewed with most universities showing between 3000 and 10000 applications for the years 1992 and 1993.
- __bowl__ - The bowl variable is a binary categorical variable. We note that over the 2 year period considered there are fewer universities that appear in bowl games than universities that do.
- __btitle__ - The btitle variable is also a binary categorical variable. The histogram, (as one would expect) shows that there are significantly fewer universities that have won a title over the 2 year period considered than universities that have.
- __finfour__ - The finfour variable is also a binary categorical variable. The histogram, (as one would expect) shows that there are significantly fewer universities that have participated in men's final four games over the 2 year period considered than universities that have.

```{r, tidy=TRUE, tidy.opts=list(width.cutoff=70)}
#How many observations and variables are in the dataset
str(data)
```

```{r, tidy=TRUE, tidy.opts=list(width.cutoff=70)}
# apps variable
print(quantile(data$apps, probs = c(.01, .05, .10, .25, .50, .75, .90, .95, .99, 1)))

# Plot the histogram of apps at 15 bins
apps.hist <- ggplot(data, aes(apps)) + 
    theme(legend.position = "none") + 
    geom_histogram(fill = "Blue", colour = "Black", binwidth = (range(data$apps)[2] - range(data$apps)[1])/15) +
    labs(title = "Distribution of Applications", x = "Number of applications for admissions", y = "Frequency")

plot(apps.hist)
```

```{r, tidy=TRUE, tidy.opts=list(width.cutoff=70)}
# bowl variable
print(quantile(data$bowl, probs = c(.01, .05, .10, .25, .50, .75, .90, .95, .99, 1)))

# Plot the histogram of bowl at 15 bins
bowl.hist <- ggplot(data, aes(bowl)) + 
    theme(legend.position = "none") + 
    geom_histogram(fill = "Blue", colour = "Black", binwidth = (range(data$bowl)[2] - range(data$bowl)[1])/15) +
    labs(title = "Distribution of Bowl Game Participation", x = "Bowl game participation", y = "Frequency")

plot(bowl.hist)
```

```{r, tidy=TRUE, tidy.opts=list(width.cutoff=70)}
# btitle variable
print(quantile(data$btitle, probs = c(.01, .05, .10, .25, .50, .75, .90, .95, .99, 1)))

# Plot the histogram of btitle at 15 bins
btitle.hist <- ggplot(data, aes(btitle)) + 
    theme(legend.position = "none") + 
    geom_histogram(fill = "Blue", colour = "Black", binwidth = (range(data$btitle)[2] - range(data$btitle)[1])/15) +
    labs(title = "Distribution of btitle variable", x = "Conf Champ title Won", y = "Frequency")

plot(btitle.hist)
```

```{r, tidy=TRUE, tidy.opts=list(width.cutoff=70)}
# finfour variable
print(quantile(data$finfour, probs = c(.01, .05, .10, .25, .50, .75, .90, .95, .99, 1)))

# Plot the histogram of finfour at 15 bins
finfour.hist <- ggplot(data, aes(finfour)) + 
    theme(legend.position = "none") + 
    geom_histogram(fill = "Blue", colour = "Black", binwidth = (range(data$finfour)[2] - range(data$finfour)[1])/15) +
    labs(title = "Distribution of finfour variable", x = "Final-Four game participation", y = "Frequency")

plot(finfour.hist)
```

Question2
============
The distribution of the change of log of application number has the appearance of a
normal(ish) distribution. There are 2 outlier points with a change of -.2 and +.4
that correspond to Arizona University and Arkansas University. There are however
no indications that these outliers would affect the regression at this point or that
they should be removed.

Which schools had the greatest increase in number of log applications?
----------------------------------------------------------------------
Arizona, Alabama and Arizona State (see table below).

Which schools had the greatest decrease in number of log applications?
----------------------------------------------------------------------
Arkansas, Oklahoma State and Penn State (see table below).

```{r, tidy=TRUE, tidy.opts=list(width.cutoff=70)}
reshaped.data <- reshape(data, v.names=c("apps","top25","ver500","mth500","stufac",
                                         "bowl","btitle","finfour","lapps",
                                         "avg500","bball","perf"),
                         timevar="year",idvar="school", direction="wide")
#Check the layout of the reshaped data
str(reshaped.data)

# Create the new variable for the change in the log of the number of applications
reshaped.data$clapps <- reshaped.data$lapps.1993-reshaped.data$lapps.1992

#examine the new variable
print(quantile(reshaped.data$clapps, probs = c(.01, .05, .10, .25, .50, .75, .90, .95, .99, 1)))

# Plot the histogram of at 20 bins
clapps.hist <- ggplot(reshaped.data, aes(clapps)) + 
    theme(legend.position = "none") + 
    geom_histogram(fill = "Blue", colour = "Black", binwidth = (range(reshaped.data$clapps)[2] - range(reshaped.data$clapps)[1])/20) +
    labs(title = "Distribution of clapps variable", x = "Change in number of log aplications", y = "Frequency")

plot(clapps.hist)
```

```{r, tidy=TRUE, tidy.opts=list(width.cutoff=70)}
# Which schools had the greatest increase in number of log applications?
head(reshaped.data[order(reshaped.data$clapps, decreasing= TRUE), c("school", "clapps")])
```

```{r, tidy=TRUE, tidy.opts=list(width.cutoff=70)}
# Which schools had the greatest decrease in number of log applications?
head(reshaped.data[order(reshaped.data$clapps, decreasing= FALSE), c("school", "clapps")])
```

Question3
==========
Which of these new variables has the highest variance?
------------------------------------------------------
The cperf variable has the highest variance, with a value of 0.82425892

```{r, tidy=TRUE, tidy.opts=list(width.cutoff=70)}
#Create the additional variables
reshaped.data$cperf <- reshaped.data$perf.1993-reshaped.data$perf.1992
print(quantile(reshaped.data$cperf, probs = c(.01, .05, .10, .25, .50, .75, .90, .95, .99, 1)))
print(stat.desc(reshaped.data$cperf))

reshaped.data$cbball <- reshaped.data$bball.1993-reshaped.data$bball.1992
print(quantile(reshaped.data$cbball, probs = c(.01, .05, .10, .25, .50, .75, .90, .95, .99, 1)))
print(stat.desc(reshaped.data$cbball))

reshaped.data$cbowl <- reshaped.data$bowl.1993-reshaped.data$bowl.1992
print(quantile(reshaped.data$cbowl, probs = c(.01, .05, .10, .25, .50, .75, .90, .95, .99, 1)))
print(stat.desc(reshaped.data$cbowl))

reshaped.data$cbtitle <- reshaped.data$btitle.1993-reshaped.data$btitle.1992
print(quantile(reshaped.data$cbtitle, probs = c(.01, .05, .10, .25, .50, .75, .90, .95, .99, 1)))
print(stat.desc(reshaped.data$cbtitle))

reshaped.data$cfinfour <- reshaped.data$finfour.1993-reshaped.data$finfour.1992
print(quantile(reshaped.data$cfinfour, probs = c(.01, .05, .10, .25, .50, .75, .90, .95, .99, 1)))
print(stat.desc(reshaped.data$cfinfour))
```

Question4
==========
a. Additional assumptions needed for the population model to be causal
----------------------------------------------------------------------
One important assumption needed for the model to be causal is the exogeneity of the dependent variables in the model with the error term in the model. Exogeneity can be defined as the requirement that the dependent variables are not related to the error term.
Assuming the population model presented:
$$lapps_{i}=\gamma_{0} + \beta_{0}I_{1993} + \beta_{1}bowl_{i} + \beta_{2}btitle_{i} + \beta_{3}finfour_{i} + a_{i} + u_{i}$$
Exogeneity of the variables in the model can be formulated as:
$$Cov(bowl_{i}, u_{i})=Cov(btitle_{i}, u_{i})=Cov(finfour_{i}, u_{i})=0$$

b.  Additional assumption needed for OLS to consistently estimate the first-difference model
--------------------------------------------------------------------------------------------
For OLS to consistently estimate the first difference model, we need to assume that assumptions MLR1-MLR4' hold. We also need to assume that the model isn't affected by omitted variable bias, meaning that all the explanatory variables are included in the model. The presence of omitted variables that are correlated with any of the variables in the model creates a bias in
the estimation of the coefficients of the variables selected in the model.

In the  abscence of a guarantee that we have not left unmeasured effect in the model, we must make the assumption that we have the ability to make manipulations to the x terms in order to observe changes in the y term, without affecting the error term, so that we can estalish that the bias introduced by any unmeasured variables is null.

This assumption is not very plausible.  It's hard to think of a situation where we can make a school appear in the final four / win a title / be champions in the previous year without affecting other variables (the error term).

Question5
===========
Model Analysis
--------------
- __F-statistic__ - The F-statistic for the model has a p-value of 0.03855, which is significant at the 0.05 level.
- __intercept__ - The coefficient for the intercept is 0.01684 indicating that the year over year incease in application is 1.6% from 1992 to 1993.  However, the t-statistic for that coefficient has a value of 0.1932 and is not significant at the 0.05 level.
- __cbowl__ - The coefficient for the cbowl variable is .057, indicating that a win in a bowl the year prior, contributes to a 5.7% increase in applications the subsequent year.  The t-statistic for the coefficient is significant at the 0.05 level, with a value of 0.0236.
- __cbtitle__ - The coefficient for the cbtitle variable is .041, indicating that a win in the men's conference championship in the previous year, translates into an increase of 4.1% year over year from 1992 to 1993.  The t-statistic for the coefficient is 0.1950 and is not significant at the 0.05 level.
- __cfinfour__ - The coefficient for the cfinfour variable is -0.06961 and seems to indicate that an appearance in the men's final four the year prior is related to a decrease of 6.9% of applications.

```{r, tidy=TRUE, tidy.opts=list(width.cutoff=70)}
#Create the model.
change.lapps.model <- lm(clapps ~ cbowl + cbtitle + cfinfour, data = reshaped.data)
#Estimate the model
print(summary(change.lapps.model))
```

Question6
==============
The F-statistic obtained from the model is the test of overall significance of the model. We have already establish that it has a p-value of 0.03855, which is significant at the 0.05 level.  We can obtain the same statistic with the linearHypothesis function.

Because the model is significant, we do not want to read too much into the t-statistics of each of the coefficients of the model. We had noted that 3 out of 4 coefficients had p-values that were not significant at the 0.05 level.  It appears that the combined explanatory power of the model is still relevant, as evidenced by the F-statistic.

```{r, tidy=TRUE, tidy.opts=list(width.cutoff=70)}
linearHypothesis(change.lapps.model, c("cbowl", "cbtitle", "cfinfour"))
```
