---
title: "W271 Lab 3 Spring 2016"
author: "Megan Jasek, Rohan Thakur, Charles Kekeh"
date: "Friday, April 22, 2016"
output: pdf_document
---

```{r,include=FALSE}
library(pastecs)
library(knitr)
library(psych)
library(astsa)
library(zoo)    
library(forecast)
library(quantmod)
library(fGarch)
library(tseries)
library(ggplot2)
library(stargazer)

opts_chunk$set(tidy.opts=list(width.cutoff=60))

# Set the seed so that reported numbers remain stable
set.seed(1)
```

```{r, tidy=TRUE, tidy.opts=list(width.cutoff=60)}
# Functions for Parts 2, 3, 4
get.best.arima <- function(x.ts, maxord = c(1,1,1))
{
  best.aic <- 1e8
  all.aics <- vector()
  all.models <- vector()
  n <- length(x.ts)
  for (p in 0:maxord[1]) for(d in 0:maxord[2]) for(q in 0:maxord[3]) 
  {
    fit <- arima(x.ts, order = c(p,d,q), method="ML")
    fit.aic <- -2 * fit$loglik + (log(n) + 1) * length(fit$coef)
    if (fit.aic < best.aic) 
    {
      best.aic <- fit.aic
      best.fit <- fit
      best.model <- c(p,d,q) 
    }
    all.aics <- c(all.aics, fit.aic)
    all.models <- c(all.models, sprintf("(%d, %d, %d)", p,d,q))
  }
  list(best=list(best.aic, best.fit, best.model), others=data.frame(aics=all.aics, models=all.models))
}

get.best.sarima <- function(x.ts, maxord=c(1,1,1,1,1,1), freq)
{
  best.aic <- 1e8
  all.aics <- vector()
  all.models <- vector()
  n <- length(x.ts)
  for (p in 0:maxord[1]) for (d in 0:maxord[2]) for (q in 0:maxord[3])
    for (P in 0:maxord[3]) for (D in 0:maxord[4]) for (Q in 0:maxord[5])
    {
      fit <- arima(x.ts, order = c(p,d,q),
                   seasonal=list(order= c(P,D,Q),freq),
                   method ="CSS", optim.control = list(maxit = 10000))
      fit.aic <- -2*fit$loglik + (log(n) + 1) * length(fit$coef)
      if (fit.aic < best.aic){
        best.aic <- fit.aic
        best.fit <- fit
        best.model <- c(p,d,q,P,D,Q)
      }
      all.aics <- c(all.aics, fit.aic)
      all.models <- c(all.models, sprintf("(%d, %d, %d, %d, %d, %d)", p,d,q,P,D,Q))
    }
  list(best=list(best.aic, best.fit, best.model), others=data.frame(aics=all.aics, models=all.models))
}

plot.time.series <- function(x.ts,bins=30, name)
{
  str(x.ts)
  par(mfrow=c(2,2))
  hist(x.ts,bins, main = paste("Histogram of", name, sep=" "),
       xlab="Values")
  plot(x.ts, main=paste("Plot of", name, sep=" "),
       ylab="Values", xlab="Time Period")
  acf(x.ts, main = paste("ACF of", name, sep=" "))
  pacf(x.ts, main = paste("PACF of", name, sep=" "))
}

plot.residuals.ts <- function(x.mod, model_name)
{
  par(mfrow=c(1,1))
  hist(x.mod$residuals,30, main=paste("Histogram of", model_name,
                                      "Residuals", sep=" "), xlab="Values")
  par(mfrow=c(2,2))
  plot(x.mod$residuals, fitted(x.mod), main=paste(model_name, "Fitted vs. Residuals", sep=" "),
       ylab="Fitted Values", xlab="Residuals")
  plot(x.mod$residuals, main=paste(model_name, "Residuals", sep=" "),
       ylab=paste("Residuals", sep=" "))
  acf(x.mod$residuals, main = paste("ACF of", model_name, sep=" "))
  pacf(x.mod$residuals, main = paste("PACF of", model_name, sep=" "))
  Box.test(x.mod$residuals, type="Ljung-Box")
}

estimate.ar <- function(x.ts)
{
  x.ar = ar(x.ts)
  print("Difference in AICs")
  print(x.ar$aic)
  print("AR parameters")
  print(x.ar$ar)
  print("AR order")
  print(x.ar$order)
  return(x.ar)
}

plot.orig.model.resid <- function(x.ts, x.mod, orig_name, model_name, xlim, ylim)
{
  df <- data.frame(cbind(x.ts,fitted(x.mod),x.mod$residuals ))
  class(df)  
  stargazer(df, type="text", title="Descriptive Stat", digits=1)

  summary(x.ts)
  summary(x.mod$residuals)
  par(mfrow=c(1,1))
  plot.ts(x.ts, col="red", 
        main=paste(orig_name, "Original vs Estimated", model_name,
                   "Series with Resdiauls", sep=" "),
        ylab=paste(orig_name, "Original and Estimated Values", sep=" "),
        xlim=xlim, ylim=ylim, pch=1, lty=2)
  par(new=T)
  plot.ts(fitted(x.mod),col="blue",axes=T,xlab="",ylab="",
        xlim=xlim, ylim=ylim, lty=1) 
  leg.txt <- c(paste(orig_name, "Original Series", sep=" "), "Estimated Series", "Residuals")
  legend("topleft", legend=leg.txt, lty=c(2,1,2), 
       col=c("red","blue","green"), bty='n', cex=1)
  par(new=T)
  plot.ts(x.mod$residuals,axes=F,xlab="",ylab="",col="green",
        xlim=xlim, ylim=ylim, lty=2, pch=1, col.axis="green")
  axis(side=4, col="green")
  mtext("Residuals", side=4, line=2,col="green")
}

plot.model.forecast <- function(x.mod, mod.fcast, orig_name, num_steps, x, y)
{
  par(mfrow=c(1,1))
  plot(mod.fcast,
       main=paste(num_steps, "-Step Ahead Forecast and", orig_name, "Original & Estimated Series", sep=" "),
       xlab="Time",
       ylab=paste(orig_name, "Original, Estimated, and Forecasted Values", sep=" "),
       xlim=x, ylim=y, lty=2, lwd=1.5)
  par(new=T)
  plot.ts(fitted(x.mod),col="blue", 
        lty=2, lwd=2, xlab="", ylab="", xlim=x, ylim=y)
  leg.txt <- c(paste(orig_name, "Original Series", sep=" "), "Estimated Series", "Forecast")
  legend("topleft", legend=leg.txt, lty=c(2,2,1), lwd=c(1,2,2),
       col=c("black","blue","blue"), bty='n', cex=1)
}

```


Part 3 (25 points): Forecast the Web Search Activity for global Warming
=======================================================================
**Data Analysis**

 1. The time series has weekly values (630 of them) starting at 1/4/04 and ending at 1/24/16.  The minimum value is -0.551 and the maximum value is 4.104.
 2. Time series plot shows that the series is very persistent,  The series is basically flat from 2004 to 2012.  After 2012, there is a sharp trend upward.  There is more volatility after 2012.  There are spikes and dips which could be seasonal with a yearly frequency.  The series is not stationary in the mean.
 3. Histogram shows is heavily positively skewed with most values between -0.551 and -0.3.
 3. ACF of the series has correlations at around 0.75 for almost 25 lags.
 4. PACF drops off immediately after first lag.  There are 4 points that fall outside the 95% confidence interval (blue lines) at lags 3, 5, 11 and 14.  The PACF could show some signs of seasonality.

**Model Selection Process**

1. __Try AR models.__  Use the ar() command in R to find AR(p) models or order p that potentially fit the time series.  This command output a model or order 15, but looking at the difference in AICs, the AIC for the AR(1) model is not that different (only 29.85 point away) from the AIC of the AR(15), so for parsimony we will try using that one.  Check if the residuals look like white noise.
- Histogram:  Yes.  This looks like a normal distribution.
- Fitted vs. Residuals:  No.  The plot does not look like an evenly distributed cloud.
- Plot:  No.  The plot does not look random, there is a lot of volatility on the right hand side of the graph.
- ACF:  No.  The ACF drops off after lag 0, but has only a few lags where the correlation comes out of the 95% confidence interval (CI)
- PACF:  No.  The PACF shows correlation with several values outside of the 95% CI.
In summary, the residuals for this model do not look like white noise, so there is more variation that could be explained by our model.  The In-Sample fit of this estimated model matches the original model very well as evidenced in the plot.

2. __Try ARIMA models.__  Use the get.best.arima() function which will try models with c(p,d,q) where p=0-4, d=0-2 and q=0-2.  And then we can print out a list in ascending order by AIC of the 20 models with the lowest AIC.  And then inspect these models for parsimony and select one with a good AIC and a small number of parameters.  The best model output from the function had an AIC of -1058.794 with parameters = c(1, 2, 2).  For parsimony a model of ARIMA(1,1,1) was chosen with an AIC of -1032.364 which is not that different from the best AIC.  Check if the residuals look like white noise.  No, the residuals do not look like white noise.  They exhibit the same characteristics as the AR(1) model from step 1.  The In-Sample fit of this estimated model matches the original model very well as evidenced in the plot.

3. __Try SARIMA models.__  From the plot of the original series, it looks like this series has a seasonal component with a 52-week periodicity.  Use the get.best.sarima() function with parameters c(2,2,2,2,2,2).  The best AIC output is -1276.817 with a model of SARIMA(1,2,2,1,0,2).  For parsimony try running get.best.sarima() with c(1,1,1,1,1,1).  A parsimonious model from this output is SARIMA(0,1,1,1,0,1) with AIC -1246.412 which is very close to the AIC output from c(2,2,2,2,2,2).  For parsimony we will choose SARIMA(0,1,1,1,0,1) and check the residuals.  No, the residuals do not look like white noise.  They exhibit the same characteristics as the AR(1) model from step 1.  The residuals plot exhibits evidence of time-varying volatility.  The In-Sample fit of this estimated model matches the original model very well as evidenced in the plot.

4. __Try using GARCH.__ Since the residuals exhibit evidence of time-varying volatility, we will try to use GARCH to model that.  A GARCH model is fit with the residuals from the SARIMA(0,1,1,1,0,1) model from step 3.  Looking at the residuals of the GARCH model, the square of the residuals is still not completely inside the 95% CI indicating that there is still time-varying volatility present.  Since we haven't found a model with a satisfactory fit, we will look at only modeling part of the original time series.

```{r, tidy=TRUE, tidy.opts=list(width.cutoff=60)}
# Read in the time series data
glob.warm = read.csv("globalWarming.csv", header=TRUE)
glob.warm.ts = ts(glob.warm$data.science, start=2004, frequency = 52)
# Print descriptive statistics
str(glob.warm.ts)
summary(glob.warm.ts)
cbind(head(glob.warm.ts), tail(glob.warm.ts))
quantile(as.numeric(glob.warm.ts), c(.01,.05,.1,.25,.5,.75,.9,.95,.99))
# Plot the time series
plot.time.series(glob.warm.ts, 50, "Global Warming")

### 1. Try AR models
glob.warm.ar = estimate.ar(glob.warm.ts)
glob.warm.ar1=arima(glob.warm.ts, order=c(1,0,0))
# Plot the residuals
plot.residuals.ts(glob.warm.ar1, "AR(1)")
# Plot the In-sample fit
plot.orig.model.resid(glob.warm.ts, glob.warm.ar1, "", "AR(1)", c(2004,2016), c(-0.7,4))

### 2. Try ARIMA models
#gw.arima.best <- get.best.arima(glob.warm.ts, maxord=c(4,2,2))
# Print the top 20 best models based on AIC
#gw.arima.best$others[order(gw.arima.best$others$aics)[1:20],]
glob.warm.arima=arima(glob.warm.ts, order=c(1,1,1))
# Plot the residuals
plot.residuals.ts(glob.warm.arima,"ARIMA(1,1,1)")
# Plot the In-sample fit
plot.orig.model.resid(glob.warm.ts, glob.warm.arima, "", "ARIMA(1,1,1)", c(2004,2016), c(-0.7,4))

### 3. Try SARIMA models
#gw.seas.best <- get.best.sarima(glob.warm.ts, maxord=c(2,2,2,2,2,2), 52)
# Print the top 20 best models based on AIC
#gw.seas.best$others[order(gw.seas.best$others$aics)[1:20],]

#gw.seas.best1 <- get.best.sarima(glob.warm.ts, maxord=c(1,1,1,1,1,1), 52)
# Print the top 20 best models based on AIC
#gw.seas.best1$others[order(gw.seas.best1$others$aics)[1:20],]
glob.warm.arima.seas = arima(glob.warm.ts, order = c(0,1,1),
                             seas = list(order = c(1,0,1), 52),
                             method="CSS")
# Plot the residuals
glob.warm.arima.seas.res = glob.warm.arima.seas$residuals
plot.residuals.ts(glob.warm.arima.seas,"SARIMA(0,1,1,1,0,1)")
par(mfrow=c(1,1))
acf(glob.warm.arima.seas.res^2)
# Plot the In-sample fit
plot.orig.model.resid(glob.warm.ts, glob.warm.arima.seas, "", "SARIMA(0,1,1,1,0,1)", c(2004,2016), c(-0.7,4))

### 4. Try using GARCH.
glob.warm.garch.fit = garchFit(~ garch(1, 1), data = glob.warm.arima.seas.res, trace = FALSE)
glob.warm.garch.res <- glob.warm.garch.fit@residuals
par(mfrow=c(2,1))
acf(glob.warm.garch.res)
acf(glob.warm.garch.res^2)
```

5. __Using a portion of the data.__ Since no satisfactory model was found using the full data series, using a portion of the data will be considered.  The data has a clear split around 2012 or 2013 where it goes from being stationary in the mean to being non-stationary in the mean.  Since we are interested in forecasting this information after 2016, we will try creating a model for the latter part of the data, the part that contains the most recent information and then forecasting after that.  2012 could have been chosen, but it still had some of the non-trending data contained in it, so 2013 was chosen as a start year.  We will repeat the same analysis as above.  When graphs are created with this 2013-2016 series, they will be denoted by the phrase "Abrv. Original" instead of "Original"

6. __Try AR models.__  Use the ar() command in R to find AR(p) models or order p that potentially fit the time series.  This command output a model or order 1.  Check if the residuals look like white noise.
- Histogram:  Yes.  This looks like a normal distribution.
- Fitted vs. Residuals:  Yes.  The plot looks like an evenly distributed cloud.
- Plot:  Yes.  The plot looks mostly like white noise.  There is a little more volatility on the right hand side of the graph.
- ACF:  No.  The ACF drops off after lag 0, but has only a few lags where the correlation comes out of the 95% confidence interval (CI)
- PACF:  No.  The PACF shows correlation with a few values outside of the 95% CI.
In summary, the residuals for this model do not look like white noise, so there is more variation that could be explained by our model.

7. __Try ARIMA models.__  Use the get.best.arima() function to find the best model.  The best model output from the function had an AIC of -25.091124 with parameters = c(0, 1, 0).  An ARIMA(0,1,0) model was created.  Check if the residuals look like white noise.  No, the residuals do not look like white noise.  They exhibit the same characteristics as the AR(1) model from step 6.  The In-Sample fit of this estimated model matches the original model very well as evidenced in the plot.

8. __Try SARIMA models.__  Use the get.best.sarima() function.  The best AIC output is -90.17105 with parameters c(0,1,1,1,0,1), but for parsimony we will choose a SARIMA(0,1,0,0,0,1) with an AIC of -76.45286 which is very close to the other model.  Check the residuals.  Yes, the residuals look basically like white noise.  There is one place in the squared residuals where the value exceeds the 95% confidence interval.  The In-Sample fit of this estimated model.  The model now has a satisfactory fit and we will move on to backtesting and forecasting.

9. __Backtesting.__ For backtesting, 10% of the values from the end of the 2013-2016 time series were withheld, in this case 10 values.  The backtesting model shows mean predicted values that follow the up and down changes of the original time series, but the mean predicted values are are not as extreme as the orignal values.  The seasonality of the original series is being modeled to some extent.  We also note that the original series, for the most part, falls within the 95% confidence interval of the forecast, giving us confidence that this model could be used as a decent predictive model for the original time series.

10. __Forecast the model.__  Using the SARIMA(0,1,0,0,0,1) model with the 2013-2016 version of the time series, we made the requested 12-step ahead forecast of the model.  The forecast looks like it captures the seasonality of the model as it matches the upward trend and the seasonal volatility.  We also note that all of the forecasted values are within the 80% confidence interval of the prediction.

**Conclusion**

MJ WILL FINISH WRITING LATE.  The 2013-2016 abbreviated time series is satisfactorily modeled with a SARMIMA(x,x,x,x,x,x) model to hande trends and seasonality.  We observe that the forecasts of the model are consisent with the seasonality of the original series and the forecasts are within the 95% confidence interval.

```{r, tidy=TRUE, tidy.opts=list(width.cutoff=60)}
### 5. Using a portion of the data
# Create 2013 to 2016 series
glob.warm.2013.ts = ts(glob.warm.ts[471:length(glob.warm.ts)],start=2013,frequency = 52)
# Create a string label to prepend to the word "Original" for proper labeling in graphs
part_name = "Abrv."
# Rename series
glob.warm.part.ts = glob.warm.2013.ts
# Descriptive statistics
str(glob.warm.part.ts)
summary(glob.warm.part.ts)
cbind(head(glob.warm.part.ts), tail(glob.warm.part.ts))
quantile(as.numeric(glob.warm.part.ts), c(.01,.05,.1,.25,.5,.75,.9,.95,.99))
# Plot the time series
plot.time.series(glob.warm.part.ts,50, "Abrv. GW 2013-2016")

### 6. Try AR models.
# Use the ar function to find an ar estimate
glob.warm.ar = estimate.ar(glob.warm.part.ts)
summary(glob.warm.ar)
# Create an AR(1) model
glob.warm.part.ar1=arima(glob.warm.part.ts, order=c(1,0,0))
# Plot residuals
plot.residuals.ts(glob.warm.part.ar1, "AR(1)")
par(mfrow=c(2,1))
# Plot the In-sample fit
plot.orig.model.resid(glob.warm.part.ts, glob.warm.part.ar1, part_name, "AR(1)", c(2013,2016), c(-0.7,4))

### Try ARIMA models
#gw.part.arima.best <- get.best.arima(glob.warm.part.ts, maxord=c(3,3,3))
# Print the top 10 best models based on AIC
#gw.part.arima.best$others[order(gw.part.arima.best$others$aics)[1:10],]

# Create ARIMA(0,1,0) model
glob.warm.part.arima=arima(glob.warm.part.ts, order=c(0,1,0))
# Plot residuals
plot.residuals.ts(glob.warm.part.arima, "AR(0,1,0)")
# Plot In-sample fit
plot.orig.model.resid(glob.warm.part.ts, glob.warm.part.arima, part_name, "AR(0,1,0)", c(2013,2016), c(-0.7,4))

### Try SARIMA models
gw.part.seas.best <- get.best.sarima(glob.warm.part.ts, maxord=c(1,1,1,1,1,1), 52)
# Print the top 20 best models based on AIC
gw.part.seas.best$others[order(gw.part.seas.best$others$aics)[1:5],]
# Create SARIMA(0,1,0,0,0,1) model
glob.warm.part.sarima = arima(glob.warm.part.ts, order = c(0,1,0),
                             seas = list(order = c(0,0,1), 52), method="CSS")
# Plot the residuals
plot.residuals.ts(glob.warm.part.sarima, "SARIMA(0,1,0,0,0,1)")
par(mfrow=c(2,1))
acf(glob.warm.part.sarima$residuals, main="Abrv. GW 2013-2016 SARIMA(0,1,0,0,0,1) Residuals")
acf(glob.warm.part.sarima$residuals^2, main="Abrv. GW 2013-2016 SARIMA(0,1,0,0,0,1) Residuals")
# Plot the In-sample fit
plot.orig.model.resid(glob.warm.part.ts, glob.warm.part.sarima, part_name, "SARIMA(0,1,0,0,0,1)", c(2014,2016.5), c(-0.7,5))

#### Backtesting
# Create a time series cutting off the last 10% of the values
glob.warm.part.bt.ts = ts(glob.warm.part.ts[1:(length(glob.warm.part.ts)-10)],start=2013,frequency = 52)
# Recreate SARIMA(0,1,0,0,0,1) model
glob.warm.part.sarima.bt = arima(glob.warm.part.bt.ts, order = c(0,1,0),
                                seas = list(order = c(0,0,1), 52), method="CSS")
# Combine the cut-off time series with the fitted values from the model and the residuals from the model
df.part = cbind(glob.warm.part.bt.ts, 
           fitted(glob.warm.part.sarima.bt),
           glob.warm.part.sarima.bt$resid)
colnames(df.part) = c("orig_series", "fitted_vals", "resid")
head(df.part)
# Plot the Abrv. Original and estimate series with residuals
par(mfrow = c(1, 1))
plot.ts(df.part[, "orig_series"], col = "red", main = "Abrv. Original vs SARIMA Estimated Series with Residuals",
        ylab = "Abrv. Original and Estimated Values", xlim = c(2013, 2016),
        ylim = c(-.7,4))
par(new = T)
plot.ts(df.part[, "fitted_vals"], col = "blue", axes = T, xlab = "",
        ylab = "", xlim = c(2013, 2016), ylim = c(-.7,4))
leg.txt <- c("Abrv. Original Series", "Estimated Series", "Residuals")
legend("top", legend = leg.txt, lty = 1, col = c("red", "navy",
                                                 "green"), bty = "n", cex = 1)
par(new = T)
plot.ts(df.part[, "resid"], axes = F, xlab = "", ylab = "", col = "green",
        xlim = c(2013, 2016), ylim = c(-.7,4), pch = 1)
axis(side = 4, col = "green")
mtext("Residuals", side = 4, line = 2, col = "green")

# Create a forecast for backtesting and plot it
glob.warm.part.sarima.bt.fcast = forecast.Arima(glob.warm.part.sarima.bt, h=10)
par(mfrow = c(1, 1))
plot(glob.warm.part.sarima.bt.fcast, lty = 2, col = "navy", main = "10-Step Out-of-Sample Forecast for Abrv. GW 2013-2016 Series",
     ylab = "Abrv. Original, Estimated, and Forecast Values",
     xlim = c(2013, 2017), ylim = c(-.7, 4.5))
par(new = T)
plot.ts(glob.warm.part.ts, axes=F, lty = 1, col = "black", xlim = c(2013, 2017),
        ylim = c(-.7, 4.5), ylab = "")
leg.txt <- c("Abrv. Original Series", "Forecast series")
legend("top", legend = leg.txt, lty = 1, col = c("black", "blue"),
       bty = "n", cex = 1)

#### Forecasting - Forecast the request 12-step ahead forecast
glob.warm.part.sarima.fcast = forecast.Arima(glob.warm.part.sarima, h=12)
print(str(glob.warm.part.sarima.fcast))
print(summary(glob.warm.part.sarima.fcast$mean))
plot.model.forecast(glob.warm.part.sarima, glob.warm.part.sarima.fcast, part_name,
                    "12", c(2013,2017), c(-0.7,5.5))

```